Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/design.tex
make spell

Nail produces the following from a single, descriptive grammmar: 
\begin{itemize}
\item type declarations for the internal model,
\item the \textit{parser}, a function to parse a sequence of bytes into an
instance of the model, and
\item the \textit{generator}, a function to create a
sequence of bytes from an instance of the model.
\end{itemize}

@@

\paragraph{Semantic Bijecction.}
Second, in order to generate output, Nail maintaining
a \emph{semantic bijection} between the external format and the internal model. That is, when
generating the corresponding output to a parser input, the two byte streams will have the same
meaning, i.e. be interpreted equivalently by Nail, but might not be identical.
 A bijection in the traditional sense often does not make sense for data
formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
that tolerates arbitrarily long padding.
%\footnote{Say, the physical layer of most communication
%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
%  pre-determined synchronization sequence and the actual contents of the transmission.}
 Program semantics should be independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded fields \emph{constants}.
Similarly, Nail does not preserve the layout of objects referred to by their offsets. If the grammar
consists of a simple protocol without offset fields, constants and the like, there is a conventional
bijection between  internal models and valid parser inputs.
\paragraph{Hide redundant information.}
Nail's internal model is designed to hide unneeded and redundant information from the application.
Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
and needs to be kept as additional state during parsing. Dependent fields are for example used to represent
lengths, offsets and checksums.

@@

In the rest of this section, we describe the design on the Nail grammar language.
\input{fig-syntax}
\subsection{Basics}
A Nail parser defines both the structure of some external format and a data type to represent that
format. Parsers are constructed by combinators over simpler parsers, an approach popularised by the Parsec
framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
 combinator libraries, such as Parsec and Hammer\cite{hammer-parser} and extend them so they also
 describe a data type.

@@

Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
approach popularised by the Hammer framework.\cite{hammer-parser}  
Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header.  The internal representation rounds up to the nearest integer
type, for example a 24-bit unsigned integer will be stored in an unsigned 32-bit double word.

@@

\paragraph{Choices.}
Our choice combinator takes multiple labelled parsers and tries each of them in turn, backtracking
to the initial position if an option fails. This behaviour for the choice combinator, the defining
feature of Parser Expression Grammars\cite{ford2002packrat}, resolves any ambiguities explicitly.
Internally, the choice combinator represents choices as tagged unions.

@@

\paragraph{Optional.}Nail also includes an \cc{optional} combinator, which attempts to recognise a value, but succeeds
without consuming input when it cannot recognise that value. Syntactically, \cc{optional} is
equivalent to a choice between the parser and an empty structure, but in the internal model it is
more concisely represented as a reference that is null when the parser fails.

@@

Dependent fields are defined within a structure like normal fields, but their name starts with an \@-sign. 
A dependent field is in scope and can be referred to by the definition of all subsequent fields in
the same structure. Dependency fields can be passed to rule invocations as parameters.

>>>>

Nail produces the following from a single, descriptive grammar: 
\begin{itemize}
\item type declarations for the internal model,
\item the \textit{parser}, a function to parse a sequence of bytes into an
instance of the model, and
\item the \textit{generator}, a function to create a
sequence of bytes from an instance of the model.
\end{itemize}

@@

\paragraph{Semantic Bijection.}
Second, in order to generate output, Nail maintaining
a \emph{semantic bijection} between the external format and the internal model. That is, when
generating the corresponding output to a parser input, the two byte streams will have the same
meaning, i.e. be interpreted equivalently by Nail, but might not be identical.
 A bijection in the traditional sense often does not make sense for data
formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
that tolerates arbitrarily long padding.
%\footnote{Say, the physical layer of most communication
%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
%  pre-determined synchronization sequence and the actual contents of the transmission.}
 Program semantics should be independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded fields \emph{constants}.
Similarly, Nail does not preserve the layout of objects referred to by their offsets. If the grammar
consists of a simple protocol without offset fields, constants and the like, there is a conventional
bijection between  internal models and valid parser inputs.
\paragraph{Hide redundant information.}
Nail's internal model is designed to hide unneeded and redundant information from the application.
Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
and needs to be kept as additional state during parsing. Dependent fields are for example used to represent
lengths, offsets and checksums.

@@

In the rest of this section, we describe the design on the Nail grammar language.
\input{fig-syntax}
\subsection{Basics}
A Nail parser defines both the structure of some external format and a data type to represent that
format. Parsers are constructed by combinators over simpler parsers, an approach popularized by the Parsec
framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
 combinator libraries, such as Parsec and Hammer\cite{hammer-parser} and extend them so they also
 describe a data type.

@@

Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
approach popularized by the Hammer framework.\cite{hammer-parser}  
Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header.  The internal representation rounds up to the nearest integer
type, for example a 24-bit unsigned integer will be stored in an unsigned 32-bit double word.

@@

\paragraph{Choices.}
Our choice combinator takes multiple labeled parsers and tries each of them in turn, backtracking
to the initial position if an option fails. This behavior for the choice combinator, the defining
feature of Parser Expression Grammars\cite{ford2002packrat}, resolves any ambiguities explicitly.
Internally, the choice combinator represents choices as tagged unions.

@@

\paragraph{Optional.}Nail also includes an \cc{optional} combinator, which attempts to recognize a value, but succeeds
without consuming input when it cannot recognize that value. Syntactically, \cc{optional} is
equivalent to a choice between the parser and an empty structure, but in the internal model it is
more concisely represented as a reference that is null when the parser fails.

@@

Dependent fields are defined within a structure like normal fields, but their name starts with an $@$-sign. 
A dependent field is in scope and can be referred to by the definition of all subsequent fields in
the same structure. Dependency fields can be passed to rule invocations as parameters.

### 2014-05-01 paper/osdi14/eval.tex
make spell

Processing these encodings in existing tools, such as Bison or Hammer, would at best be very
awkward, because some ad-hoc trick would have to be used to re-position the parsers input stream.
Keeping track of the position all recognised labels would not be enough, as the offset field may
refer to any byte within the packet, not just the beginning of labels. For this reason, the DNS
server used as a tutorial example for Hammer does not support compression.

@@

Our grammar is also very efficient to use. Using our grammar, we build a ZIP file extractor in 50
lines of C. Because more recent versions of ZIP have added more features, such as large file support
and encryption, the closest existing tool in functionality is the historic version 5.4 of the Info-Zip unzip
utility\cite{infozip} that is shipped with most Linux distributions. The entire unzip distribution
is about 46.000 lines of code, which is mostly optimised implementation of various compression
algorithms and other configuration and portability code.

@@

\subsection{Security.} 
We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
analyse a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
server.

@@

Finally, we used the \cc{queryperf} tool provided with BIND to query each domain between zero
and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 
core  of an Intel i7-3610QM with 12GB of RAM on an idle system, and the benchmark tool kept at most 20
queries outstanding at once. The queryperf tool was configured to
repeat the same, randomised sequence of queries for one minute and the average throughput and
response latency were measured.  We repeated each test seven times  with 50.000
domain names, restarting each daemon in between. We also performed one initial dry run to warm the
file system cache for the zone file. We repeated the tests with 1 million domain names and found
that the servers performed almost identically.

Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
of optimisation by a factor of 3.  This demonstrates that even though Nail was
not developed focusing on performance, applications built on Nail can compete with real-world
hand-written parsers.

>>>>

Processing these encodings in existing tools, such as Bison or Hammer, would at best be very
awkward, because some ad-hoc trick would have to be used to re-position the parsers input stream.
Keeping track of the position all recognized labels would not be enough, as the offset field may
refer to any byte within the packet, not just the beginning of labels. For this reason, the DNS
server used as a tutorial example for Hammer does not support compression.

@@

Our grammar is also very efficient to use. Using our grammar, we build a ZIP file extractor in 50
lines of C. Because more recent versions of ZIP have added more features, such as large file support
and encryption, the closest existing tool in functionality is the historic version 5.4 of the Info-Zip unzip
utility\cite{infozip} that is shipped with most Linux distributions. The entire unzip distribution
is about 46.000 lines of code, which is mostly optimized implementation of various compression
algorithms and other configuration and portability code.

@@

\subsection{Security.} 
We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
analyze a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
server.

@@

Finally, we used the \cc{queryperf} tool provided with BIND to query each domain between zero
and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 
core  of an Intel i7-3610QM with 12GB of RAM on an idle system, and the benchmark tool kept at most 20
queries outstanding at once. The queryperf tool was configured to
repeat the same, randomized sequence of queries for one minute and the average throughput and
response latency were measured.  We repeated each test seven times  with 50.000
domain names, restarting each daemon in between. We also performed one initial dry run to warm the
file system cache for the zone file. We repeated the tests with 1 million domain names and found
that the servers performed almost identically.

Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
of optimization by a factor of 3.  This demonstrates that even though Nail was
not developed focusing on performance, applications built on Nail can compete with real-world
hand-written parsers.

### 2014-05-01 paper/osdi14/relwk.tex
make spell

Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
generating output for Ocaml.  Parsifal structures grammars as an Ocaml
type that holds an internal model and functions for parsing input and
output.  However, their generator can produce parsers and generators
only for simple, fixed-size structures.  The programmer can then use
these when implementing parsers and generators for more complicated
formats, manually handling offsets, checksums, and the like, risking bugs.
Nail handles more complicated constructs without the programmer manually
writing code to support them.

>>>>

Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
generating output for OCaml.  Parsifal structures grammars as an OCaml
type that holds an internal model and functions for parsing input and
output.  However, their generator can produce parsers and generators
only for simple, fixed-size structures.  The programmer can then use
these when implementing parsers and generators for more complicated
formats, manually handling offsets, checksums, and the like, risking bugs.
Nail handles more complicated constructs without the programmer manually
writing code to support them.
